%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\section{Thesis Structure}
Chapter~\ref{Background} gives an overview of MDE by defining terminology; describing associated engineering principles, practices and tools; and identifying related areas of computer science. Section~\ref{sec:mde_benefits_and_challenges} synthesises some of the benefits of -- and challenges to -- contemporary MDE.

Chapter~\ref{LiteratureReview} reviews theoretical and practical software evolution research. Areas of research that underpin software evolution are described, including refactoring, design patterns, and traceability. The review then discusses work that approaches particular categories of evolution problem, such as programming language, schema and grammar evolution. Section~\ref{subsec:mde_evo} surveys work that considers the evolution of MDE development artefacts (such as models and transformations). Section~\ref{sec:literature_review_summary} identifies three types of evolution that occur in MDE projects and highlights challenges for their management.

Chapter~\ref{Analysis} surveys existing MDE projects and categories the types of evolution occurring in each project. From this survey, a context for the thesis research is established (a type of evolution termed model-metamodel co-evolution). Examples of evolution taken from the existing MDE projects are used to identify the strengths and weaknesses of existing structures and processes for identifying and managing evolution. From this, Section~\ref{subsec:user-driven_co-evolution} identifies a process for identifying and managing evolution which has not been recognised previously in the literature, Section~\ref{subsec:co-evolution_categorisation} derives a categorisation of the existing structures and processes, and Section~\ref{sec:requirements_identification} synthesis requirements for novel structures and processes for identifying and managing evolution.

Chapter~\ref{Implementation} describes three novel structures and process for identifying and managing evolution in the context of MDE. The first, a metamodel-independent syntax, is used to identify -- and to facilitate the reconciliation of -- problems caused by evolution. The textual modelling notation described in Section~\ref{sec:notation} and the model migration language described in Section~\ref{sec:flock} are used for reconciliation of problems caused by evolution. The latter provides a means for performing reconciliation in a repeatable manner.

Chapter~\ref{Evaluation} assesses the novel structures and processes developed in Chapter~\ref{Implementation} by comparison to existing structures and processes. To explore the research hypothesis, several different types of comparison were performed, including an experiment in which quantitive measurements were derived, a collaborative comparison of model migration tools with three MDE experts, and application to a large, independent example of evolution taken from a real-world MDE project.

Chapter~\ref{Conclusion} summarises the achievements of the research, and discusses results in the context of the research hypothesis. Limitations of the thesis research and areas of future work are also outlined.