%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\declaration{
	Except where stated, all of the work contained in this thesis represents the original contribution of the author. Section~\ref{sec:collaborative_comparison} reports collaborative experiments with model migration tools, and that section makes clear the roles of the author and other participants.

	Parts of the work described in this thesis have been previously published by the author in:

	\begin{itemize}
		\item \textbf{The Epsilon Generation Language}, Louis M. Rose, Ri\-ch\-ard F. Paige, Dimitrios S. Kolovos and Fiona A.C. Polack in \textit{Proc. European Conference on Model Driven Architecture -- Foundations and Applications (ECMDA-FA)}, volume 5095 of LNCS, pages 1-16. Springer, 2008, \\ \cite{rose08egl}.
	
		\item \textbf{Constructing Models with the Human-Usable Textual Notation}, Louis M. Rose, Richard F. Paige, Dimitrios S. Kolovos and Fiona A.C. Polack in \textit{Proc. International Conference on Model Driven Engineering Languages and Systems (MoDELS)}, volume 5301 of LNCS, pages 249-263. Springer, 2008, \cite{rose08hutn}.
	
		\item \textbf{An Analysis of Approaches to Model Migration}, Louis M. Rose, Richard F. Paige, Dimitrios S. Kolovos and Fiona A.C. Polack in \textit{Proc. Joint Model-Driven Software Evolution and Model Co-evolution and Consistency Management (MoDSE-MCCM) Workshop}, co-located with MoDELS 2009, \cite{rose09analysis}.
	
		\item \textbf{Enhanced Automation for Managing Model and Metamodel Inconsistency}, Louis M. Rose, Dimitrios S. Kolovos, Richard F. Paige and Fiona A.C. Polack in \textit{Proc. International Conference on Automated Software Engineering (ASE)}, pages 545-549, ACM Press, 2009, \cite{rose09enhanced}.
	
			\item \textbf{Concordance: An Efficient Framework for Managing Model Integrity}, Louis M. Rose, Dimitrios S. Kolovos, Nicholas Drivalos, James. R. Williams, Richard F. Paige, Fiona A.C. Polack, and Kiran J. Fernandes in \textit{Proc. European Conference on Modelling Foundations and Applications (ECMFA)}, volume 6138 of LNCS, pages 62-73. Springer, 2010, \cite{rose10concordance}.

		\item \textbf{Model Migration with Epsilon Flock}, Louis M. Rose, Dimitrios S. Kolovos, Richard F. Paige, and Fiona A.C. Polack in \textit{Proc. International Conference on the Theory and Practice of Model Transformations (ICMT)}, volume 6142 of LNCS, pages 184-198. Springer, 2010, \cite{rose10flock}.
	
		\item \textbf{Model Migration Case}, Louis M. Rose, Dimitrios S. Ko\-lo\-vos, Richard F. Paige, and Fiona A.C. Polack in \textit{Proc. Transformation Tools Contest (TTC)}, co-located with TOOLS Europe 2010, \cite{rose10ttc_case}.
	
		\item \textbf{Migrating Activity Diagrams with Epsilon Flock}, \\ Louis M. Rose, Dimitrios S. Kolovos, Richard F. Paige, and Fiona A.C. Polack in \textit{Proc. Transformation Tools Contest (TTC)}, co-located with TOOLS Europe 2010, \\ \cite{rose10ttc_solution}.
		
		\item \textbf{A Comparison of Model Migration Tools}, Louis M. Rose, Markus Herrmannsdoerfer, James R. Williams, Dimitrios S. Kolovos, Kelly Garc\'{e}s, Richard F. Paige, and Fiona A.C. Polack in \textit{Proc. International Conference on Model Driven Engineering Languages and Systems (MoDELS)}, volume 6394 of LNCS, pages 61–75. Springer, 2010, \\
		\cite{rose10comparison}.
	\end{itemize}
}