%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\chapter{Evaluation}
\label{Evaluation}
This chapter evaluates the structures and processes proposed in this thesis with respect to the research hypothesis outlined in Section~\ref{sec:hypothesis}. In particular, the evaluation explores the extent to which the structures and processes affect developer productivity and the understandability of systems. As discussed in Chapter~\ref{Introduction}, 
these two characteristics are key to the success of the large and complicated software engineering projects demanded by society today. Given the time constraints of a doctoral thesis, the proposed structures and processes are prototypes, and unlikely to be fit for industrial use in their current state. The evaluation aims to determine areas in which the proposed structures and processes might be usefully improved by identifying factors that affect the efficacy of the structures and processes. The conclusions drawn in this chapter identify benefits and drawbacks of using the proposed structures and processes for managing co-evolution in contemporary MDE development environments and for industrial software-engineering projects.

Chapter~\ref{Analysis} identified \emph{user-driven co-evolution}, a process for managing co-evolution, that had been used in real-world MDE projects and that had not been recognised in the literature. Chapter~\ref{Implementation} described the implementation of two structures tailored for user-driven co-evolution, a \emph{metamodel-independent syntax} and a \emph{textual modelling notation}. Using a real-world example of user-driven co-evolution, Section~\ref{sec:exemplar_user-driven_co-evo} assesses the extent to which the dedicated structures proposed in Chapter~\ref{Implementation} affect the productivity of user-driven co-evolution.

The remainder of the chapter evaluates developer-driven co-evolution (in which a migration strategy is specified in an executable format) and focuses on \emph{Epsilon Flock} (Section~\ref{sec:flock}), a transformation language tailored for model migration. Section~\ref{sec:quantitive} evaluates the novel source-target relationship strategy implemented in Flock, \emph{conservative copy}, by comparison with two existing source-target relationship strategies using co-evolution examples from real-world projects. Sections~\ref{sec:collaborative_comparison} and~\ref{sec:ttc} evaluate Flock as a whole, using an expert evaluation and a transformation contest, respectively. The transformation contest was judged by members of the MDE community, providing an opportunity for Flock and other transformation tools to be assessed in a peer review.

The work presented in this chapter has been published in \cite{rose10comparison,rose10ttc_solution,rose10ttc_case}. The evaluation described in Sections~\ref{sec:collaborative_comparison} and \ref{sec:ttc} was performed collaboratively, and the contributions of others are highlighted in those sections.

\input{6.Evaluation/user_driven_example.tex}
\input{6.Evaluation/quantitive.tex}
\input{6.Evaluation/collaborative_comparison.tex}
\input{6.Evaluation/ttc.tex}
\input{6.Evaluation/limitations.tex}


\section{Chapter Summary}
The work presented in this chapter assessed the structures and processes proposed in Chapter~\ref{Implementation}. The evaluation has investigated the extent to which the prototypical implementations of the structures and processes increase developer productivity and the understandability of systems. Factors that affect the efficacy of the proposed structures and processes were identified via the evaluation, and are summarised below. A range of techniques were used for evaluation, and the work presented in this chapter provides a foundation for future work on -- and further evaluation of -- structures and processes for managing and identifying model-metamodel co-evolution.

Section~\ref{sec:exemplar_user-driven_co-evo} explored and compared the ways in which user-driven co-evolution is performed with and without two of the structures proposed in Chapter~\ref{Implementation}. The comparison suggested that the way in which models are represented and the way in which conformance is checked affects developer productivity. A model representation that is optimised for human rather than machine use might reduce the likelihood of mistakes, particularly if the developers performing co-evolution lack expertise in manipulating XMI (the canonical model storage representation used by contemporary modelling frameworks). Checking conformance as the model is migrated might prevent errors, particularly if several types of breaking metamodel change are to be resolved, but might be a waste of computational resources when few breaking metamodel changes are to be resolved.

Section~\ref{sec:quantitive} investigated the conciseness of model migration strategies written using new-target, existing-target and conservative copy source-target relationships. The results indicate that conservative copy yields the most concise model migration strategies, which might lead to increased developer productivity when specifying migration manually. However, other factors that were not assessed in this chapter will affect productivity, such as the learnability of the three source-target relationships. When specifying migration with an operator-based or an inference approach, migration strategies are typically presented to the user to facilitate selection between feasible alternatives. Further evaluation is needed to assess whether conciseness of migration strategies is a desirable quality for operator-based and inference approaches.
 
Sections~\ref{sec:collaborative_comparison} and~\ref{sec:ttc} assessed Epsilon Flock (Section~\ref{sec:flock}), a model-to-model transformation language tailored for model migration, by comparison with developer-driven co-evolution tools and model-to-model transformation languages. The comparisons illuminated various considerations for co-evolution. For example, the requirements and maturity of the MDE process can preclude the use of some co-evolution approaches. When an established MDE process is used to develop a system, it might be impossible to use co-evolution approaches that mandate specific tools or techniques, such as operator-based approaches, which require that evolution is expressed with a specialised metamodel editor.

In addition to the evaluation described in this chapter, the work presented in this thesis has been subject to peer review by the academic and Eclipse communities. The thesis research has been published in 3 workshop papers, 2 European conference papers and 4 international conferences papers. Epsilon HUTN, Epsilon Flock and Concordance (Chapter~\ref{Implementation}) were contributed to the Epsilon project, a component of the research incubator for the Eclipse Modeling Project (EMP), which is arguably the most active MDE community at present. EMP's research incubator hosts a limited number of participants, selected through a rigorous process and contributions made to the incubator undergo regular technical review. 

Evaluation of the research hypothesis has highlighted areas in which the structures and processes proposed in Chapter~\ref{Implementation} should be improved, and has motivated plans for future work, which are described in Section~\ref{sec:future_work}. Additionally, the work presented in this chapter has identified areas in which further evaluation might be carried out. In particular, comprehensive user studies are needed to assess the usability and learnability of co-evolution tools, and to further evidence the benefits of co-evolution tools in terms of increased developer productivity and understandability of systems.