%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\subsection{Transformation Tools Contest}
The Transformation Tools Contest (TTC) is a workshop series that seeks to compare and contrast tools for performing model and graph transformation. At TTC 2010, two rounds of submissions were invited: cases (transformation problems, three of which are selected by the workshop organisers) and solutions to the selected cases. In addition, TTC 2010 include a \emph{live contest}: during the workshop a further transformation problem was announced and solutions submitted.

Participation in TTC 2010 facilitated further evaluation of Flock. Flock and XX transformation tools were assessed for a model migration problem based on a real-world example of metamodel evolution from the UML \cite{uml21}. As part of the live contest, Flock was also assessed along with XX transformation tools for a model transformation problem. Compared to the evaluation described in Section~\ref{subsec:collaborative}, the evaluation in this section compares Flock to a wider range of tools (model and graph transformation tools, and not just model migration tools), and investigates the suitability of Flock for specifying model transformation.

The remainder of this section describes the model migration problem (Section~\ref{}) and Flock solution (Section~\ref{}), and the use of Flock for specifying a model transformation in the live contest.

\subsubsection{Model Migration Case}

\subsubsection{Model Migration Solution in Epsilon Flock}

\subsubsection{Epsilon Flock in the Live Contest}


\subsubsection{Summary}
- Summarise work

- Contributions
-- Established a real-world example of migration that can be used in further comparison / evaluation work.
-- Flock won two awards.

- Threats to validity
-- I chose the migration case. Some mitigation: based on real world example, discussion on the forums helped to identify ambiguities. 

