%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\chapter{Conclusions}
\label{Conclusion}
% The conclusion will provide a summary of the challenges addressed by, and the objectives of, our research. We will summarise the way in which we have approached the challenges and met the objectives, concluding with a summary of the evaluation and discussion of future work.

This thesis has investigated software evolution -- a key and costly development activity in software engineering \cite{moad90maintaining} -- in the context of Model-Driven Engineering (MDE), a state-of-the art approach to software engineering. While MDE promises increased developer productivity \cite{watson08mdahistory} and increased portability of software systems \cite{frankel02mda}, it also poses several challenges that threaten its adoption. For \cc example, identifying and managing evolutionary change in the context of MDE presents many open research challenges \cite{Mens07}. The thesis research has contributed to these research challenges and has explored the following research hypothesis:

\begin{quote}
\emph{In existing MDE projects, the evolution of MDE development artefacts is typically managed in an ad-hoc manner with little regard for re-use. Dedicated structures and processes for managing evolutionary change can be designed by analysing evolution in existing MDE projects. Furthermore, supporting those dedicated structures and processes in contemporary MDE environments is beneficial in terms of increased productivity for software development activities pertaining to the management of evolutionary change.}
\end{quote} 

To explore the thesis hypothesis, the following research objectives were defined.

\begin{enumerate}
	\item Identify and analyse the evolution of MDE development artefacts in existing projects.
	\item Investigate the extent to which existing structures and processes can be used to manage the evolution of MDE development artefacts. 
	\item Propose and develop new structures and processes for managing the evolution of MDE development artefacts, and integrate those structures and processes with a contemporary MDE development environment.
	\item Evaluate the proposed structures and processes for managing evolutionary change, particularly with respect to productivity.
\end{enumerate}

The remainder of this chapter summarises the contributions of the thesis in relation to the thesis hypothesis and research objectives, and gives a brief description of and motivation for several potential extensions to the thesis research.


\input{7.Conclusion/contributions}
\input{7.Conclusion/future_work}


\section{Closing Remarks}
Building the systems demanded by society now and in the future will require new approaches to software engineering \cite{selic03pragmatics}. MDE is a state-of-the-art, principled approach to software engineering, and promises many benefits particularly with respect to the portability and maintainability of software systems \cite{kleppe03mda,frankel02mda}. While MDE shows promise, its success is reliant on the availability of mature and powerful tools. Such tools are beginning to emerge, but typically fail to address concerns that affect their applicability to the engineering of large and complex software systems, such as scalability and the cost of systems evolution.

The work presented in this thesis has demonstrated a systematic method for identifying challenges for software evolution in typical MDE processes, proposed structures and processes for addressing those challenges, and evaluated the structures and processes by comparison to related work and by application to real-world examples of evolution.