%!TEX root = /Users/louis/Documents/PhD/Deliverables/Thesis/thesis.tex

\section{Limitations of the Proposed Structures and Processes}
\label{sec:limitations}
The limitations of the research presented in the thesis are now discussed. Some of the shortcomings identified here are elaborated on in Section~\ref{sec:future_work}, which highlights areas of future work.

\paragraph{Generality} The thesis research focuses on model-metamodel co-evolution, but, as discussed in Chapter~\ref{Analysis}, metamodel changes can affect artefacts other than models. Model management operations and model editors are specified using metamodel concepts and, consequently, are affected when a metamodel changes. The work presented in Chapter~\ref{Implementation} focuses on migrating models in response to metamodel changes, and does not consider integration with tools for migrating model management operations and model editors. To reduce the effort required to manage the effects of metamodel changes, it seems reasonable to envisage a unified approach that migrates models, model management operations, model editors, and other affected artefacts.

\paragraph{Reproducibility} The analysis and evaluation presented in Chapters~\ref{Analysis} and \ref{Evaluation} respectively involved using migration tools to understand and assess their functionality. With the exceptions noted below, the work presented in these chapters is difficult to reproduce and therefore the results drawn are somewhat subjective. On the other hand, multiple approaches to analysis and evaluation have been taken, and the work has been published and subjected to peer review. 

Not all of the work in Chapter~\ref{Analysis} and \ref{Evaluation} is difficult to reproduce. In particular, the evaluation methods used in Chapter~\ref{Evaluation} are described in detail and a complete set of results are provided in Appendices~\ref{CoevolutionExamples} and~\ref{TTCScores}. In general, the lack of real-world examples of co-evolution restricts the extent to which any work in this area can be considered reproducible. 

\paragraph{Formal semantics} No formal semantics for the conservative copy algorithm (Section~\ref{sec:flock}) have been provided. Instead, a reference implementation, Epsilon Flock, was developed, which facilitated comparison with other migration and transformation tools. Without a reference implementation, the evaluation described in Sections~\ref{sec:quantitive}, \ref{sec:collaborative_comparison} and~\ref{sec:ttc} would have been impossible. For \cc Epsilon as a whole, a similar case has been made in favour of a reference implementation over a formal semantics \cite{kolovos09thesis}. For domains where completeness and correctness are a primary concern, a formal semantics would be required before Flock could be applied to manage model-metamodel co-evolution.  